---
title: "Week 4 Lab"
author: "Madison DiGirolamo"
date: "9/18/2024"
output: html_document
---


```{r}
library(knitr)
opts_knit$set(root.dir = "C:/Users/madid/OneDrive/Documents/Desktop/RCour
                     seExamples/R-Class-2024/WK04_Bootstrapping")


```

Goals of Lab 4: + review SD, SE, confidence intervals + for() loops + build boostrap function

```{r}
x <- rnorm(1000)
#modeling with normal distribution and 1000 observations
hist(x, freq= FALSE, # makes the y-axis density
    ylim = c(0,2), #y-axis scale
    xlim = c(-4,4),#x-axis scale
    main = "1000 observations") #title


mean.x <- replicate(1000, mean(rnorm(1000)))
#find the mean of the distribution, 1000 times
#add the distribution of th emean to the histogram
hist(mean.x, add =TRUE,
     col = "blue", #color of plot to blue
     freq = FALSE) #to make y-axis density


```

# Checkpoint 1 (1000 obs)

-   what is the region where 95% of the data points lie?

```{r}
s1 <- sd(x) #SD
s1
a1 <- c(-1,1)*1.96*s1 + mean(x) 
a1 #95% CI for where the data points lie

```

what is the region where there is a 95% chance the mean of the population exists?

```{r}
SE1 <- (s1/sqrt(1000))
SE1

b1 <- mean(x) + c(-1,1)*1.96*SE1
b1 #95% CI for where the mean of the population exists

```

The region for the 95% chance of where the mean exists is smaller than the 95% region for where all of the data exists. Looking back at the histogram, area in blue is much narrower than the data

Recreate the same plots but with only 25 observations.

```{r}
x <- rnorm(25)
hist(x,
     freq = FALSE,
     ylim = c(0,2),
     xlim = c(-4,4),
     main = "25 observations")

mean.x <- replicate(1000, mean(rnorm(25)))
hist(mean.x, add = TRUE, # adds to the preexisting plot
     col = "blue",
     freq = FALSE)
```

# Checkpoint 2

What is the region where 95% of the data points lie? This is a larger region than with 1000 observations.

```{r}
s2 <- sd(x)
s2
a2 <- c(-1,1)*1.96*s2 + mean(x)
a2

```

What is the region where there is a 95% chance that the mean of the population exists (25 observations)? This is also less narrow than with 1000 observations because with less observations we can not be as certain.

```{r}
SE2 <- (s2/sqrt(25))
SE2

b2 <- mean(x) + c(-1,1)*1.96*SE2
b2

```

End of Checkpoint 2:

Some of this is answered above but for both the 95% CI for the data and the 95% CI for the mean, the CI is larger when there are less observations.

# Creating For Loop Function

```{r}
My.CI <- function(x.vect){
# x.vect should be normally distributed
sd.x <- sd(x.vect)
n <- length(x.vect)
se.x <- sd.x/sqrt(n)
data.95.CI <- mean(x.vect)+1.96*c(-1,1)*sd.x

mean.95.CI <- mean(x.vect)+1.96*c(-1,1)*se.x

return(list(sd=sd.x, se=se.x, mean= mean(x.vect),
data.95 = data.95.CI,
mean.95 = mean.95.CI))
}

My.CI(x) #where x is 25 observations

```

The function above assumes normal distribution, but what do you when your data
does not fit a normal distribution? You can bootstrap, which resamples the data
to create a sampling distribution.

```{r}
my.output <- c()#empty vector to store output
my.input <- seq(53, 128, by=5) #sequence we will use to calculate output

for (i in 1:length(my.input)){my.output[i] <- my.input[i]^2 + my.input[i]}

my.output

```

```{r}
my.input

```

```{r}
sample(my.input, replace = TRUE) #samples the vector with replacement

```

My.input and the sample form my.input with replacement are not the same.

```{r}
# I already used s1 so here I use s10 instead
#resample the two groups 100 times with different probabilities (prob line)
s10 <- sample(c("Group 1", "Group 2"), 
              size = 100, 
              replace = TRUE,
              prob = c(0.3,0.7))
table(s10)

```

For the next example we are using fish growth data-- fish growth rate (mm/year) as a function of temp (C)

```{r}
#reading csv file
fish.df <- read.csv("FishGrowthTemp.csv", header=TRUE)
#creating a plot of the data
par(mfrow=c(1,1))
plot(fish.df$temp, fish.df$growth.rate,
     xlab ="Temp (C)",
     ylab = "Growth Rate (mm/y)")
abline(lm(fish.df$growth.rate~fish.df$temp)) #adds line

```

```{r}
#checking if the data looks like it follows a normal distribution using both
#a histogram and a qqplot

#histogram
hist(lm(fish.df$growth.rate~fish.df$temp)$residuals,
     xlab = "Growth Rate (mm/y)")

#qqplot
qqnorm(fish.df$growth.rate)
qqline(fish.df$growth.rate)
```

```{r}
lm(fish.df$growth.rate~fish.df$temp)$coef
#checking for intercept and slope
```

```{r}
summary(lm(fish.df$growth.rate~ fish.df$temp))

```

# Checkpoint 3

Although it looks like there is a positive relationship based on the qqplot, it
does not seem to follow a normal distribution and so the t-tests done with the
summary function should not be trusted, even though we still could find
a significant relationship/slope greater than 0 using another distribution or a 
bootstrap.

Using a bootstrap to see if slope is different than 0 by resampling 100 times.

```{r}
n.boot <- 1000

#fill with
slope <- rep(NA, n.boot)
n.rows <- nrow(fish.df)
#create slopes
for (i in 1:n.boot){
  new.rows <- sample(1:n.rows, replace =TRUE)
  new.data <- fish.df[new.rows,]
  slope[i] <- lm(new.data$growth.rate ~ new.data$temp)$coef[2]
}

#histogram of slopes
hist(slope, breaks = 100)

```

```{r}
sort(slope)[n.boot*0.025] #lower 95% CI
sort(slope)[n.boot*0.975] #upper95% CI
```

# Checkpoint 4 
When using the bootstrap method, we see it is possible still that the slope 
could be zero. It is far to one side of the CI, but still possible. Adding more 
bootstraps in theory increases the power of the test, but the result would 
be negligible here because we are already having it do such a high amount. 
Looking at SE in a bootstrap example would be useless because SE is based on 
the normal distribution, and bootstraps are used when the data does not fit
a normal distribution, but you can look at a 95% CI of the mean. 


# Really Non-normal Example

```{r}
n = 100000
set.seed(1)
x1 <- rpois(n/2, 2)
x2 <- rpois(n/2, 19)

data <- c(x1, -x2)
hist(data, xlab = "x", main = "Really Non-normal data")
```

The data is very obviously not normal.

```{r}
mean(data) #mean of the data, not sample mean
```

```{r}
#sample mean
set.seed(1)
my.sample <- sample(data, 30, replace = TRUE)
mean(my.sample)

```

```{r}
se <- sd(my.sample)/sqrt(length(my.sample))
mean.95 <- mean(my.sample) + 1.96*c(-1,1)*se
print(mean.95)

```

The 95% confidence interval includes the true mean.


Sampling Distribution of the Mean
```{r}
n.boot <- 10000
mean.est <- rep(NA, n.boot)

for (i in 1:n.boot){
  my.sample <- sample(data,30, replace = TRUE)
  mean.est[i] <- mean(my.sample)
}

```

```{r}
hist(mean.est, xlab = "Sample Means")
sort(mean.est)[n.boot*0.025] #lower
sort(mean.est)[n.boot*0.975] #upper CI on distribution
abline(v = c(sort(mean.est)[n.boot*0.025], sort(mean.est)[n.boot*0.975]),
             lty =2,
             col = "red")

```
Even for highly non-normal data, the sampling distribution of the mean is 
normal!
