---
title: "report"
author: "Madison DiGirolamo"
date: '2024-09-15'
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "C:/Users/madid/OneDrive/Documents/Desktop/RCour
                     seExamples/R-Class-2024/WK03_Reproducibility")
```

## Creating the directory structure by adding folders

```{r}
if (!file.exists('data')) dir.create('data') #creates the file because 
#it does not exist in the location
if (!file.exists('src')) dir.create('src')
if (!file.exists('analysis')) dir.create('analysis')
if (!file.exists('figures')) dir.create('figures')
if (!file.exists('report')) dir.create('report')
```

I created a script called "driver.R" in the src folder and a second
script called "EditDataframe.R" in src.

# Code from EditDataFrame.R Script

```{r}
#creates the data frame by reading the csv file
setwd("C:/Users/madid/OneDrive/Documents/Desktop/RCourseExamples/R-Class-2024/WK03_Reproducibility")
read.csv("data/tempExperiment-raw.csv")

df.t1 <- read.csv("data/tempExperiment-raw.csv")

#separate temperature treatment from population using strsplit
#the unlist and matrix functions rearrange the output of strsplit
a <- strsplit(as.character(df.t1$temp), split = '-') 
#as.character indicates that it is a character
newvar <- matrix(unlist(a), ncol =2, byrow = TRUE) 
# takes apart the temp and population and creates a matrix of two colums
head(newvar)# view the new table created


#creating new data frame
df.t2 <- df.t1
df.t2$temp <- newvar[,1]
#assigns column 1 of the newvar matrix to the temp column on the new data frame
df.t2$pop <- newvar[,2]
# assigns column 2 of the newvar matrix to a new population column

head(df.t2) #view new df

#now I have the correct columns that I want, but within those columns the
#formats differ in temp it is numeric (10) or the word ten, pop is pop or
#population
#need to make them uniform because they are different levels

df.t2$temp <- sub('ten', '10', df.t2$temp)
df.t2$temp <- sub('twenty', '20', df.t2$temp)
df.t2$pop <- sub('population', 'pop', df.t2$pop)
df.t2$pop <- sub('pop 1','pop1', df.t2$pop)
df.t2$pop <- sub('pop2','pop 2', df.t2$pop)
df.t2$pop <- sub('population 1','pop 1', df.t2$pop)
str(df.t2) #look at structure, found temp to be a character and population to 
#be a character
# need to change these to be factors instead of characters 



df.t2$temp <- factor(df.t2$temp) #changes temp to be a factor
df.t2$pop <- factor(df.t2$pop) #changes pop to be a factor
str(df.t2)#looking at new structure to confirm chr changed to factor
write.csv(df.t2, 'data/tempExperiment_V2.csv') # save new df as a csv


```

## Checkpoint 1

In the temperature column there are multiple formats that were used to
indicate which population each sample is from. This is not ready for
analysis because these need to be uniform to actually allow the system
to identify each sample as the right population.

## Checkpoint 2

I was able to fix all of the issues with levels in the temp and pop
columns.

## Creating the Driver

At this step I created a new script called driver.R. All code included
in the driver is below under "Driver Code"

## Checkpoint 3

We separated the editing of the data frame to be a separate script to
streamline the driver. This means that if someone opens just the driver
they would need to run everything that happened in the editing of the
data frame. The "source" line at the beginning of driver.R does this,
simultaneously streamlining the driver script but ensuring
repeatability.

# Code for Analysis script

I created a new script called Analysis, stored in src, that holds all of
the figures and ANOVA analysis.

```{r}

setwd("C:/Users/madid/OneDrive/Documents/Desktop/RCourseExamples/R-Class-2024/WK03_Reproducibility")
source("src/EditDataframe.R")
#clearing list
rm(list = ls(all= TRUE))
#reads in CSV of edited data frame
df.t <- read.csv("data/tempExperiment_V2.csv")
str(df.t) #investigate structure

#I need to change temperature and population to be factors
df.t$temp <- factor(df.t$temp)
df.t$pop <- factor(df.t$pop)

#how does temp, population, and size affect growth rate?
#visualizing the data

boxplot(df.t$growthRate ~ df.t$temp + df.t$pop) 

#boxplot version 2
boxplot(df.t$growthRate ~ df.t$temp + df.t$pop,
        names = c('10', '20', '10', '20'), #renaming the labels
        at = c(1,2,4,5),
        ylab = 'Growth rate mm/day',
        xlab = '')

#using mtext to add margin text
mtext('Pop 1', side = 1, at =1.5, line =3)
mtext( 'Pop 2', side = 1, at = 4.5, line =3)

pdf('figures/MyBoxplot.pdf', width = 5, height = 5)

boxplot(df.t$growthRate ~ df.t$temp + df.t$pop,
        names = c('10', '20', '10', '20'), #Temperature label
        at = c(1, 2, 4, 5), # space out the boxes according to population
        ylab = 'Growth rate mm/day',
        xlab = ''
)
# Introducing the 'mtext' command to add margin text
mtext('Pop 1', side = 1, at = 1.5, line = 3)
mtext('Pop 2', side = 1, at = 4.5, line = 3)
dev.off()

#ANOVA analysis
m1 <- aov(df.t$growthRate ~ df.t$temp * df.t$pop)
m1.summary <- summary(m1)
m1.summary #view the summary

saveRDS(m1.summary, 'analysis/m1.summary.rds') #saves summary of ANOVA test to 
#analysis folder


```

## Checkpoint 4

Each test is checking how the factor impacts growth rate. Neither temp
or population significantly impacts growth rate. Additionally, temp and
population combined does not significantly impact growth rate. ( All
P-Values are high)

# Driver Code

```{r}
setwd("C:/Users/madid/OneDrive/Documents/Desktop/RCourseExamples/R-Class-2024/WK03_Reproducibility")


source("src/EditDataframe.R") #this reads in the information and data frames 
    #created in EditDataframe.R

source("src/Analysis.R") # reads in the analysis script


```

## Checkpoint 5

The organization system and steps we used is easy to follow and
reproduce because the driver shows the chronology and all steps are
annotated and included in the other scripts. If someone opened this
project, along with the associated initial CSV file, they could produce
the exact same statistical analysis and figures.

## Checkpoint 6

This markdown file will knit to pdf.

## Checkpoint 7

I changed a data point in the EditDataFrame section of a new markdown and ran
the whole exact markdown again successfully. The analysis produced slightly 
different numbers. The alternate markdown is called report7.rmd in the report 
folder with this document. 


## Checkpoint 8

I have committed all changes and pushed them to the remote repository through
Github Desktop and checked in Github.com.
